---
title: "Recent Earthquakes Around Italy"
author: "Alessandro Vullo"
date: "20 October 2017"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Overview

- Aim: create and display an interactive map of recent earthquakes around Italy and the surrounding seas
- Data taken from the [Italian National Institute of Geophysics and Volcanology](http://www.ingv.it/en/)

## Data Processing

- Data available at the [INGV page](http://info.terremoti.ingv.it/) is taken using the [INGV webservices](http://webservices.ingv.it) and downloaded as a text file to a local file.

```{r, eval=FALSE, echo = TRUE}
if(!file.exists('earthquakes.txt')) {
  download.file('http://webservices.ingv.it/fdsnws/event/1/query?starttime=2017-10-09T00%3A00%3A00&endtime=2017-10-16T23%3A59%3A59&minmag=2&maxmag=10&mindepth=-10&maxdepth=1000&minlat=-90&maxlat=90&minlon=-180&maxlon=180&minversion=100&orderby=time-asc&format=text&limit=10000', destfile = 'earthquakes.txt', method = "curl")
}
```

```{r getdata}
if(!file.exists('earthquakes.txt')) {
  download.file('http://webservices.ingv.it/fdsnws/event/1/query?starttime=2017-10-09T00%3A00%3A00&endtime=2017-10-16T23%3A59%3A59&minmag=2&maxmag=10&mindepth=-10&maxdepth=1000&minlat=-90&maxlat=90&minlon=-180&maxlon=180&minversion=100&orderby=time-asc&format=text&limit=10000', destfile = 'earthquakes.txt', method = "curl")
}
```

- The downloaded file is a simple text with header and columns delimited by '|'.

```{r, eval=FALSE,echo=TRUE}
df <- read.csv('earthquakes.txt', header=TRUE, sep="|")
```

```{r loaddata}
df <- read.csv('earthquakes.txt', header=TRUE, sep="|")
```

## Data Processing (2)

```{r, eval = FALSE, echo = TRUE}
str(df)
```

```{r}
str(df)
```

## Filtering Data

- Dataset contains data for places outside Italy and the surrounding sea

```{r, eval = FALSE, echo = TRUE}
subset(df, !(Longitude>=6.0 & Longitude<=18.5 & 
            Latitude>=35.0 & Latitude<=48.0), 
            select=c('Latitude','Longitude','EventLocationName'))
```

```{r}
subset(df, !(Longitude>=6.0 & Longitude<=18.5 & Latitude>=35.0 & Latitude<=48.0), 
       select=c('Latitude','Longitude','EventLocationName'))
```

- These observations are filtered out

```{r, eval = FALSE, echo = TRUE}
df <- subset(df, Latitude < 0)
```

```{r filter}
df <- subset(df, Latitude > 0)
```

## Generating the Map

- Use [plotly](https://plot.ly/r/) to make a scatter plot of Lat. vs Long
- Color associated to the depth (Km) of the earthquake
- Size of each point proportional to the intensity

```{r map, eval = FALSE, echo = TRUE, message=FALSE, warning=FALSE}
library(plotly)
with(df, plot_ly(x = Longitude, y = Latitude, 
                 type="scatter", mode="markers", 
                 color = ~Depth.Km, size=Magnitude) %>% 
                 layout(title = "Earthquakes Around Italy", 
                        xaxis = list(title="Longitude"), 
                        yaxis = list(title = "Latitude")))
```

##

```{r plot, message=FALSE, warning=FALSE}
library(plotly)
with(df, plot_ly(x = Longitude, y = Latitude, 
                 type="scatter", mode="markers", 
                 color = ~Depth.Km, size=Magnitude) %>% 
                 layout(title = "Earthquakes Around Italy", 
                        xaxis = list(title="Longitude"), 
                        yaxis = list(title = "Latitude")))
```
